A meta-analysis of fungicide spray schedules to minimise yield loss from powdery mildew in Australian mungbean

### 2020-08-31 03_Prep_data.Rmd
Rename file

## Subset to Trials containing DMI fungicides

@@

```{r Var_mean_sqaures}
PM_MB_dat %>%
   ggplot(aes(x = grain_yield.t.ha., y = yield_error)) +
   geom_point()
```
No association between yield and sample variance, this might be due to the unpredictable nature of mungbeans.

@@

```{r Var_MeanSquarePlot}
PM_MB_dat %>%
   ggplot(aes(x = grain_yield.t.ha., y = Y_Msquare/replicates)) +
   geom_point()
PM_MB_dat[is.na(PM_MB_dat$Y_Msquare),]

@@

@@

First we need to calculate the time between the first sign of disease and the fungicide treatments.

@@

To ensure sufficient number of replicates, we will bin the continuous variable into three categorical variables relating to when the first fungicide application was made.

@@

   - **Early**: First fungicide application was prior to first sign of disease.

   - **Recommended**: First fungicide application was applied on the day powdery mildew was observed, or within three days of first sign.

   - **Late**: First fungicide application was four or more days after first sign of disease being observed.

@@

```{r}
data.frame(TreatmentName = c("Early", "Recommended", "Late", "EarlyPlus", "RecommendedPlus", "LatePlus"),
           n_sprays = rep(c("Single", "Two - Three"), each = 3),
           DaysRelativeToFirstSign = c("Prior to first sign of Powdery Mildew",
                                       "1 - 3 days after first sign of Powdery Mildew",
                                       "7 - 8 days after first sign of Powdery Mildew"))%>%
   kable()
```

@@

@@

# Perhaps move below to post analysis, analysis

### Mungbeans age and yield saved

The time at which powdery mildew infects in the seasons and the age of the plant might affect the AUDPC and the loss in yield.
Let's have a look at the age of each trial when powdery mildew infected it.
We will use the difference between `planting_date` and `first_sign_disease` as a proxy for crop maturity when it became infected.

```{r age_andYield}
dat1 <- PM_MB_dat %>%
   mutate(days2PM = first_sign_disease - planting_date) %>%
   arrange(desc(days2PM)) %>%
   select(
      trial_ref,
      year,
      location,
      spray_management,
      days2PM,
      grain_yield.t.ha,
      yield_gain,
      prop_yield_gain,
      AUDPC_m,
      D_pres,
      PM_final_severity
   )
dat1
```

From these trials I am most interested in the experiments that directly compare `Recommended` with `Recommended_P_plus` and/or `Late` with `Late_plus`. 
Therefore I will not inspect, `AM1304` (Marys Mount), which only trialled an early, prophylactic spray.

We can see that the crops which were likely the most mature, did not show large yield gains, in particular:

   * Hermitage 2019  
   * Premer  
   * Goolhi  
   * Millmerran

Let's look at this in a quick box plot. 
We will only look at trials where powdery mildew set in 60 days after planting, when the physiological maturity was likely flowering.

```{r ageBoxPlot}
dat1 %>%
   filter(days2PM >= 60) %>%
   filter(spray_management != "control" &
             trial_ref != "AM1304") %>%
   group_by(trial_ref, spray_management) %>%
   ggplot(aes(x = spray_management, y = yield_gain, colour = trial_ref)) +
   geom_boxplot() +
   scale_fill_usq() +
   geom_hline(yintercept = 0)
```

We can see in these experiments five experiments there is clearly no benefit to applying a second fungicide application given the values of the recommended _plus treatment were predominantly lower than zero.

For completeness let's test this with powdery mildew infection commencing 40 - 60 days after planting, instead of greater than 60.

```{r PM40_2_59_days}
dat1 %>%
   filter(days2PM < 60 &
             days2PM >= 40) %>%
   filter(spray_management != "control") %>%
   group_by(trial_ref, spray_management) %>%
   ggplot(aes(x = spray_management, y = yield_gain, colour = trial_ref)) +
   geom_boxplot() +
   scale_fill_usq() +
   geom_hline(yintercept = 0)
```

We can see here all four experiments benefited, on average, from more than one spray.

Let's estimate the age of the crop based on crop maturity observations, we will use the `PM_MB_dat` which include some observations of crop maturity.

```{r whatAreTheAges}
dat2 <- PM_MB_dat %>%
   mutate(
      flower = lubridate::ymd(flowering_date) - lubridate::ymd(planting_date),
      pod = lubridate::ymd(pod_fill_date) - lubridate::ymd(planting_date),
      late_pod = lubridate::ymd(mid_late_pod_fill) - lubridate::ymd(planting_date)
   ) %>%
   select(location, year, flower, pod, late_pod) %>%
   distinct() %>%
   arrange(desc(flower))

index1 <-
   apply(dat2, 1 , function(X)
      ! all(is.na(X[c("flower", "pod", "late_pod")])))

dat2[index1,]
as.double(mean(dat2[index1, "flower"], na.rm = TRUE))
as.double(mean(dat2[index1, "pod"], na.rm = TRUE))
as.double(mean(dat2[index1, "late_pod"], na.rm = TRUE))
```

We can see that flowering and podding can be variable, and depend on the cumulative degree days.
However, on average flowering starts `r as.double(mean(dat2[index1,"flower"], na.rm = TRUE))` days after sowing, podding starts `r as.double(mean(dat2[index1,"pod"], na.rm = TRUE))`, and late podding `r as.double(mean(dat2[index1,"late_pod"], na.rm = TRUE))`.

>>>>

## Subset to trials containing DMI fungicides

@@

```{r Var_mean_sqaures}
PM_MB_dat %>%
   ggplot(aes(x = grain_yield.t.ha., y = yield_error)) +
   geom_point()
```

No association between yield and sample variance, this might be due to the unpredictable nature of mungbeans.

@@

```{r Var_MeanSquarePlot}
PM_MB_dat %>%
   ggplot(aes(x = grain_yield.t.ha., y = Y_Msquare/replicates)) +
   geom_point()

@@

There is a somewhat better trend here, if sampling variance from standard deviation proves unsuitable for the analysis we may compare models using variance from stdev and mean square for the best fit.

*******

@@

First are going to calculate the time between the first sign of disease and the fungicide applications.

@@

To ensure sufficient number of replicates, we will bin the continuous variable into three categorical variables relating to when the first fungicide application was made, relative to the first sign of disease.

@@

   - **Early**: First fungicide application was prior to first sign of disease. 
   - **Recommended**: First fungicide application was applied on the day powdery mildew was observed, or within three days of first sign. 
   - **Late**: First fungicide application was four or more days after first sign of disease being observed.

@@

```{r TreatmentTable}
data.frame(TreatmentName = c("Early", "Recommended", "Late", "EarlyPlus", "RecommendedPlus", "LatePlus"),
           n_sprays = rep(c("Single", "Two - Three"), each = 3),
           DaysRelativeToFirstSign = c("Prior to first sign of Powdery Mildew",
                                       "1 - 3 days after first sign of Powdery Mildew",
                                       "7 - 8 days after first sign of Powdery Mildew"))%>%
   kable()
```

@@

******

@@

```{r saveData_PrepD}
save(PM_MB_dat, file = here("cache/PrepData.Rdata"))
```
